---
title: "Assignment 4"
author: "Pratheek Sreerangam"
date: "2022-11-02"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading the Required packages

```{r}
library(flexclust)
library(cluster)
library(tidyverse)
library(factoextra) 
library(FactoMineR)
library(ggcorrplot)
```

a.Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on.

## Loading the data

```{r}
getwd()
setwd("C:/Users/prath/Downloads")
Info<- read.csv("C:/Users/prath/Downloads/Pharmaceuticals.csv")
## Choosing columns 3 to 11 now, and putting the information in variable Info 1
Info1 <- Info[3:11]
## Displaying the top six rows of data with the head function
head(Info1)
summary(Info1)
```

##The data in Info1 and the Info updated dataframe will be scaled according to the varying weights assigned to each variable along the rows. using the factoextra package's get dist and fviz dist functions to measure the distance between data rows and visualize the distance matrix

```{r}
Infoupdated <- scale(Info1)
row.names(Infoupdated) <- Info[,1]
distance <- get_dist(Infoupdated)
fviz_dist(distance)
```

## To check the correlation between key variables, create a correlation matrix and print it.

```{r}
corr <- cor(Infoupdated)
ggcorrplot(corr, outline.color = "grey50", lab = TRUE, hc.order = TRUE, type = "full") 
```

##The ROA, ROE, Net Profit Margin, and Market Cap are all high, according to the Correlation Matrix.. 

## Finding out the relative importance of the primary variables in the data set will be done using principal component analysis. 
## assuming the optimal cluster size is 5.
```{r}
pca <- PCA(Infoupdated)
var <- get_pca_var(pca)
fviz_pca_var(pca, col.var="contrib",
             gradient.cols = c("grey","yellow","purple","red","blue"),ggrepel = TRUE ) + labs( title = "PCA Variable Variance")
```

## Using the elbow technique to discover the ideal number of customers, we can infer from PCA Variable Variance that ROA, ROE, Net Profit Margin, Market Cap, and Asset Turnover contribute over 61% to the two PCA components/dimensions Variables.
```{r}
set.seed(10)
wss <- vector()
for(i in 1:10) wss[i] <- sum(kmeans(Infoupdated,i)$withinss)
plot(1:10, wss , type = "b" , main = paste('Cluster of Companies') , xlab = "Number of Clusters", ylab="wss")
wss
```

## Exactly as predicted, the ideal cluster is at number 5.

## Silhouette Approach
## determining the optimal cluster size.

```{r}
fviz_nbclust(Infoupdated, kmeans, method = "silhouette")
```

## This demonstrates that five clusters are the optimum number. Using the k-means method to create a 5 cluster.
```{r}
set.seed(1)
k5 <- kmeans(Infoupdated, centers = 5, nstart = 25) # k = 5, number of restarts = 25
k5$centers 
k5$size  
fviz_cluster(k5, data = Infoupdated) 
```

## Manhattan Distance when using Kmeans Clustering.

```{r}
set.seed(1)
k51 = kcca(Infoupdated, k=5, kccaFamily("kmedians"))
k51
#Using predict function.
clusters_index <- predict(k51)
dist(k51@centers)
image(k51)
points(Infoupdated, col=clusters_index, pch=19, cex=0.3)
```

b.Interpret the clusters with respect to the numerical variables used in forming the clusters Using Kmeans method to calculate Mean.

```{r}
Info1 %>% mutate(Cluster = k5$cluster) %>% group_by(Cluster) %>% summarise_all("mean")
clusplot(Infoupdated,k5$cluster, main="Clusters",color = TRUE, labels = 2,lines = 0)
```

Companies are categorized into different clusters as follows:

Cluster 1: ELN, MRX, WPI and AVE
Cluster 2: AGN and PHA
Cluster 3: AHM,WYE,BMY,AZN, LLY, ABT, NVS and SGP
Cluster 4: BAY, CHTT and IVX
Cluster 5: JNJ, MRK, PFE and GSK

From the means of the cluster variables, it can be derived as follow:

Cluster 1has the best Net Profit Margin, the lowest PE ratio, and the fastest sales growth. It can be bought or kept on hand as a reserve.
 
Cluster 2 PE ratio is very high 

Cluster 3 has a medium risk

Cluster 4 Despite having an excellent PE ratio, it is incredibly risky to own due to its extremely high risk, extremely high leverage, and poor Net Profit margin. Also very low is revenue growth.

Cluster 5 has strong market capitalization, ROI, ROA, ROA on assets, ROA on turnover of assets, and ROA on net profit margin. A low PE ratio indicates that the stock price is moderately valued and may thus be bought and kept. Revenue growth of 18.5% is also favorable.

c.Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used informing the clusters)

## examining patterns by visualizing clusters against the variables

```{r}
Info3 <- Info[12:14] %>% mutate(Clusters=k5$cluster)
ggplot(Info3, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters')
ggplot(Info3, mapping = aes(factor(Clusters),fill = Location))+geom_bar(position = 'dodge')+labs(x ='Clusters')
ggplot(Info3, mapping = aes(factor(Clusters),fill = Exchange))+geom_bar(position = 'dodge')+labs(x ='Clusters')
```

## The variable and clusters There is a trend in the median recommendations.

## There doesn't seem to be any discernable pattern among the clusters, locations, or exchanges other than the fact that the majority of the clusters/companies are listed on the NYSE and situated in the United States.

d.Provide an appropriate name for each cluster using any or all of the variables in the data set.

Cluster 1: Top Buying
Cluster 2: Significant Risk
Cluster 3: Attempt it 
Cluster 4: Very Dangerous or Runaway 
Cluster 5: A Perfect Asset 

